Embedding Intuitionistic into Classical Logic

### 2022-07-12 CSL/lipics-v2021-EIICL.tex
keywords
IQC to CQC, Conclusion

\keywords{Intuitionistic Logic, Automated Reasoning, Proof Theory} %TODO mandatory; please add comma-separated list of keywords

@@

\begin{theorem}
	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$. For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$. Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal S^\#$ by including formulas as follows:
	\begin{itemize}
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to(A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item  $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
	\end{itemize}
	Then $\bigwedge S\to P$ is intuitionistically iff $\mathcal E\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
	\begin{align*}
		\mathcal E := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
		& \bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
		& \bigwedge\{\forall\vec z\forall u(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is an $n$-ary function symbol in original signature}\}
	\end{align*}
	The size of the obtained formula is linear in the input. However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
\end{theorem}

@@

@@

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{k}$ inductively as follows:
	\begin{itemize}
		\item $A^{k} = A(k)$ for every propositional variable $A$.
		\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
		\item $(\varphi\vee\psi)^k = \varphi^k\vee\psi^k$.
		\item $(\varphi\to \psi)^k = \forall u(k\preceq u\to\varphi^{u}\to\psi^{u})$ where $u$ is some new free variable.
	\end{itemize}
	Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$
	\begin{align*}
		K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
	where $u$ is some free variable.
\end{definition}

@@

\section{Quantifiers in Classical Logic}

Of course translating a propositional formula to a first-order formula is somewhat unsatisfying. As a first step to obtaining a propositional formula we wish to eliminate the quantifiers. Since we are interested in preserving validity we use Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

Now usually in classical logic before removing quantifiers prenexification is performed, however since we wish to apply a similar procedure to intuitionistic logic later, where prenexification is not valid, we will do without.

which follows from the following two Lemmata, proofs of which can be found in the appendix.

@@

Now trying to apply Herbrandization to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated. We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form. A similar translation was proposed in~\cite{statman1979intuitionistic}.

@@

Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
$$\hspace{-0.8cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$

@@

\begin{definition}
	Suppose $\mathcal S$ is a set of formulas of the form
	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition. Obtain $\mathcal S^\#$ by including the formulas
	\begin{itemize}
		\item $A^k\to A^{k\psi}$ for $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
	\end{itemize}
\end{definition}

@@

It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much. Furthermore we will be able to give fragments of IPC for which validity is complete for other stages of the polynomial hierarchy. The translation from the last section will serve as a useful tool for this. We will not explain the semantics of QBF in depth. Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$\hspace*{-.7cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and a formula $$\varphi = \bigwedge \mathcal S\to P$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

>>>>

\keywords{Intuitionistic Logic, Automated Reasoning, Propositional Logic, First-order Logic, QBF} %TODO mandatory; please add comma-separated list of keywords

@@

\begin{theorem}\label{fullFOtranslation}
	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$. For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$. Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal S^\#$ by including formulas as follows:
	\begin{itemize}
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to(A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
		\item  $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
	\end{itemize}
	Then $\bigwedge S\to P$ is intuitionistically iff $K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
	\begin{align*}
		K := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
		& \bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
		& \bigwedge\{\forall\vec z\forall u(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is an $n$-ary function symbol in original signature}\}
	\end{align*}
	The size of the obtained formula is linear in the input. However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
\end{theorem}

@@

\section{Preliminaries}

Before proceeding towards the results let us quickly fix our notation and recapitulate the semantics for intuitionistic logic.

\subsection{Propositional Semantics}

\begin{definition}
	A \textbf{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textbf{truth value} $v(A)\in\{0, 1\}$. We extend $v$ to a function $v^*$ on formulas:
	\begin{itemize}
		\item $v^*(\bot) = 0$
		\item $v^*(A) = v(A)$ for each propositional variable $A$.
		\item $v^*(\varphi\wedge\psi) = \min(v^*(\varphi), v^*(\psi))$
		\item $v^*(\varphi\vee\psi) = \max(v^*(\varphi), v^*(\psi))$
		\item $v^*(\varphi\to \psi) = \max(1 - v^*(\varphi), v^*(\psi))$
	\end{itemize}
	$v$ is a \textbf{model} for $\varphi$ if $v(\varphi) = 1$. We denote the set of valid formulas with CPC.
\end{definition}

One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation, i.e. $\neg\varphi$ is not satisfiable. The same does not hold true for intuitionistic logic as we shall see.

\begin{definition}
	A \textbf{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \textbf{persistency condition}. We extend the valuations $v_u$ slightly differently to $v_u^\#$ than before, i.e. have
	\begin{itemize}
		\item $v_u^\#(\bot) = 0$
		\item $v_u^\#(\varphi) = v_u(A)$ if $\varphi$ is atomic.
		\item $v_u^\#(\varphi\wedge\psi) = \min(v_u^\#(\varphi), v_u^\#(\psi))$
		\item $v_u^\#(\varphi\vee\psi) = \max(v_u^\#(\varphi), v_u^\#(\psi))$
		\item $v_u^\#(\varphi\to \psi) = \min\{\max(1 - v_w^\#(\varphi), v_w^\#(\psi))\:|\:w\geq u\}$
	\end{itemize}
	That is $v_u^\#(\varphi\to \psi) = 1$ if at every world $w\geq u$ we have $v_w^\#(\varphi) = 0$ or $v_w^\#(\psi) = 1$. We say that $\varphi$ is satisfied at a node $u$ if $v_u^\#(\varphi) = 1$ and write $u\models\varphi$. If a formula is satisfied at every node then $\mathcal K$ is a \textbf{model} for $\varphi$. We denote the set of valid formulas with IPC.
\end{definition}
There are many classical tautologies which are not intuitionistically valid
\begin{example}\label{LEMcounterexample}
	Consider the famous law of excluded middle $A\vee\neg A$. Recall that $\neg A := A\to \bot$. As a counter-model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$. Then $u\not\models A$ and $u\not\models \neg A$, i.e. $u\not\models A\vee\neg A$. 
\end{example}
In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable. This is insufficient for intuitionistic logic, e.g. $A\vee\neg A$ is not a intuitionistic tautology but its negation is not satisfiable either.

\subsection{Predicate Semantics}

We now give account of the more involved semantics of first-order logic

\begin{definition}
	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the domain of $\mathcal{M}$, and a function $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
		m&\text{if $b=a$}\\
		v(b)&\text{otherwise}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$.
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. We denote the set of valid formulas with CQC.
\end{definition}

\begin{definition}
	A $\Sigma$-\textit{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
	\begin{itemize}
		\item $M_u\subseteq M_w$.
		\item $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
		\item $R^{I_w}|_{M_u} = R^{I_u}$.
	\end{itemize}
	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
	\begin{itemize}
		\item $u, v\not\models\bot$
		\item $u, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I_u, v},\dots,t_n^{I_u, v})\in R^{I_u}$.
		\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$.
		\item $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. We denote the set of valid formulas with IQC.
\end{definition}
In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically.
Consider for instance the classical tautology $\varphi$ $$\neg\forall x A(x)\to \exists x \neg A(x)$$Analogous remarks with regard to validity and satisfiability hold as in the propositional case.

\subsection{Skolemization and Herbrandization}
An important step in the embedding will be the elimination of quantifiers via Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

which follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix. We are now ready to present our embedding of intuitionistic into classical logic.

@@

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively as follows:
	\begin{itemize}
		\item $A^{u} = A(u)$ for every propositional variable $A$.
		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
	\end{itemize}
	Let $K(\varphi)$ encode the Kripke semantics, i.e.
	$$K(\varphi) = \text{PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{Persistent}(u, w))$$
	with e.g.
	\begin{align*}
		\text{PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		Persistent(u, w)=&\:\bigwedge \{A(u)\to A(w))\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
	\end{align*}
	Then define
	$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
	where $u$ is some free variable.
\end{definition}

@@

@@

Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying. As a first step we want to apply Herbrandization. Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated. We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form. A similar translation was proposed in~\cite{statman1979intuitionistic}.

@@

Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$

@@

\begin{definition}
	Suppose $\mathcal S$ is a set of formulas $\psi$ of the form
	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi),(A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition. Obtain $\mathcal S^\#$ by including the formulas
	\begin{itemize}
		\item $A^k\to A^{k\psi}$ for each $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
		\item $A^k_\psi\to (B^k_\psi\circ C^k_\psi)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^k_\psi\circ B^k_\psi)\to C^k_\psi$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^{k\psi}_\psi\to B^{k\psi}_\psi)\to C^k_\psi$ for $(A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
	\end{itemize}
\end{definition}

@@

It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much. The translation from the last section will serve as a useful tool for this. We will not explain the semantics of QBF in depth. Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and$$\varphi = \bigwedge \mathcal S\to P$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$. Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < k$ as well as the special cases
	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
	for leafs and
	and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas would be
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
	\end{align*}
\end{definition}
